---
title: "hw3_real"
author: "Elizabeth Braun"
date: "2024-03-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(ggplot2)
```

## Intro

This is an analysis of the power required for different cars and car
speeds on a highway. We're adding a new "super light" car, and computing
the mean and max power using FOR loops and the pmap function.

## Power function

This function is going to calculate the amount of power needed to move a
car at whichever given speed

```{r setup, include=FALSE}
power <- function(V, A, m) {
  # Constants
  rho <- 1.225 # kg/m^3, density of air
  Cd <- 0.3 # Drag coefficient, assuming a common value for simplicity
  
  # Convert speed from km/hr to m/s
  V_m_s <- V * 1000 / 3600 
  
  # Power calculation
  P <- 0.5 * rho * Cd * A * V_m_s^3 + m * 9.81 * V_m_s # Simplified power equation
  return(P)
}
```

## Cars

Here is where we define the specifications for each car

```{r}
possible_cars <- data.frame(
  name = c("A", "B", "C", "Super Light"),
  mass = c(20000, 65000, 38000, 5000), # in kg
  area = c(22, 30, 22, 10) # in m^2
)
```

## Speeds

Here is where we generate speeds. The mean highway speed is 80 km/hr,
the other speeds will be centered around that amount with a standard
deviation of 10km/hr.

```{r}
set.seed(123) #random number generator for generating speeds
nsample <- 200
mean_speed <- 80 # km/hr
sd_speed <- 10 # km/hr

speed_samples <- rnorm(n = nsample, mean = mean_speed, sd = sd_speed)
```

## FOR loop

The FOR loop: -initializes variables and generates 100 sample speeds
based on the mean and sd, in m/s -loops through each car type defined
above through each of the 100 sample speeds, calculates their power
using the power function, and stores the resulting values -calculates
the mean and max power from the power values for each car -stores all
these calculations in the results data frame

```{r}
# Generate speeds in m/s
set.seed(123) #same as above and ensures reproducibility
nsample = 100
mean_speed_kmh = 80
sd_speed_kmh = 10
mean_speed_ms = mean_speed_kmh / 3.6 #convert from km/h to m/s
sd_speed_ms = sd_speed_kmh / 3.6 #convert from km/h to m/s

#initialize data frames to store results
results <- data.frame(car = character(), mean_power = numeric(), max_power = numeric(), stringsAsFactors = FALSE)

powers <- as.data.frame(matrix(ncol=4, nrow=nsample))
colnames(powers) = possible_cars$name

#loop through each car type
for (i in 1:nrow(possible_cars)) {
  car <- possible_cars$name[i]
  speeds <- rnorm(nsample, mean = mean_speed_ms, sd = sd_speed_ms)

  
    powers[,i] <- autopower(V = speeds, A = possible_cars$area[i], m = possible_cars$mass[i])
    
  
  mean_power <- mean(powers[,i])
  max_power <- max(powers[,i])
  
  results <- rbind(results, data.frame(car = car, mean_power = mean_power, max_power = max_power))
}

#pivot powers data frame to be longer 
powers_pivot <- pivot_longer(powers,cols = A:'Super Light',names_to = "cars",values_to = "powers")


results
```

## pmap

Here we use the pmap function from the purrr package to get the mean and
max power for each car type across the different sampled speeds. First
we generate a vector of speeds, converted from km/hr to m/s:

```{r}
# Assuming a mean speed of 80 km/hr and standard deviation of 10 km/hr
mean_speed_kmh <- 80
sd_speed_kmh <- 10
nsample <- 100

# Convert mean and sd to m/s
mean_speed_ms <- mean_speed_kmh / 3.6
sd_speed_ms <- sd_speed_kmh / 3.6

# Generate speeds
set.seed(123) # for reproducibility
speeds_ms <- rnorm(nsample, mean = mean_speed_ms, sd = sd_speed_ms)
```

Here we used pmap to compute the mean and max power for each car type
(while assuming the possible_cars dataframe was already defined):

```{r}
autopower <- function(V, A, m) {
  g <- 9.8 # m/s^2
  rho <- 1.225 # kg/m^3
  cdrag <- 0.3 # coefficient of drag
  Pb <- (0.5 * rho * A * cdrag * V^3) + (m * g * V)
  return(Pb)
}

# Function to compute mean and max power for a given car and speeds
compute_powers <- function(A, m, speeds) {
  powers <- pmap_dbl(list(V = speeds), autopower, A = A, m = m)
  return(list(powers=powers))
}


# Apply compute_powers for each car
results_pmap <- map2_dfc(possible_cars$area, possible_cars$mass, ~compute_powers(.x, .y, speeds_ms))
colnames(results_pmap) = possible_cars$name


##Pivot data frame longer
powers_pivot_pmap <- pivot_longer(results_pmap,cols = A:'Super Light',names_to = "cars",values_to = "powers")

mean_power_pmap <-apply(results_pmap, 2, mean)
max_power_pmap <- apply(results_pmap, 2, max)

mean_power_pmap
max_power_pmap

```

## Plot the data

```{r}

ggplot(powers_pivot, aes(x = cars, y = powers, fill = cars)) +
  geom_boxplot() +
  labs(title = "Power Distribution for Different Types of Cars",
       x = "Car Type",
       y = "Power (Watts)") +
  theme_minimal()
```

## Plot the pmap Data

```{r}

ggplot(powers_pivot_pmap, aes(x = cars, y = powers, fill = cars)) +
  geom_boxplot() +
  labs(title = "Power Distribution for Different Types of Cars",
       x = "Car Type",
       y = "Power (Watts)") +
  theme_minimal()
```
